---
title: 5. Calculate metrics on the shape of an extreme event."
author: "Victoria Delannoy, Nicolas Loiseau, Sébastien Villéger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Calculate metrics on the shape of an extreme event.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Aims of the function *BEE.calc.metrics_morpho()*.
Compute daily metrics of spatial features of the extreme event (only in surface, 2D).

# Usage
## Arguments:
**corrected_rasters** :
  The SpatRaster you want to analyse (first output of 
  *BEE.id.extreme_events()*).
**start_date** :
  Allows to perform the analysis on a specific  time_frame, this allows to 
  save computation time.
**end_date** :
  Allows to perform the analysis on a specific  time_frame, this allows to 
  save computation time.
**per_pix** :
  Use TRUE if you want a list with one dt per pixel as an output.
**crs** :
  To get accurate length and area, data in longitude latitude must
  be converted into meters data that take in account earth sphericity, the
  conversion depend on studdied region, thus you must specify the one the most
  appropriated for your data. To find a suitable EPSG you can
  check on https://spatialreference.org/explorer.html or https://epsg.io/ or build your own EPSG here https://projectionwizard.org/.


## Warnings:
You will receive a warning if :

-   *"start_date"* and/or *"end_date"* formats do not match the format of the spatrasters layer name. Please note that some mismatches in the date format may still not be detected. To be sure, compare your *"start_date"* and *"end_sate"* with terra::names(corrected_rasters[[1:10]]).

## Messages:
You will receive a message in the following case :  
-    The number of NA pixels is not consitent through the layers, some metrics

## Output
It is a list of dataframe (one per gps point). Each dataframe contains information on the periods of extreme events and the periods in between extreme events. Information can be daily or grouped by time periods (extreme events or in between extreme events). To simplify the explanation of the columns, the term 'period' is used to refer to both periods of extreme events and the periods between extreme events. To check if a period is extreme, you can use cleaned_value == 1 or check the daily_category column.  
Categories are defined in Hobday *et al.* 2018.  

Dataframes contains the folowing informations:

|   Column name        |             Description              |   Class   |  Unit          |
|:--------------------:|:------------------------------------:|:---------:|:--------------:|
| pixel_id             | Pixel number in the spatraster       | integer   |                |
| original_value       | Day more extreme than baseline = 1   | double    |                |
|                      | Day less extreme than baseline = 0   |           |                |
| cleaned_value        | Day belong to an exteme event = 1    | double    |                |
|                      | Day outside of an extreme event = 0  |           |                |
| event_id             | Number of the extreme event in the   | integer   |                | 
|                      | time serie                           |           |                |
| duration             | Duration of the period               | integer   | days           |  
| date                 | Date associated to the observation   |double/date|                |  
| ID                   | The event identifier \*              | character |                |
| value                | Value of the sudied parameter        | double    | \*             |
| x                    | Longitude                            | double    | decimal degree |
| y                    | Latitude                             | double    | decimal degree |
| first_date           | First day of the period              | Date      | \*\*           |
| last_date            | Last day of the period               | Date      | \*\*           |
| mean_value           | Mean value of the period             | double    | \*             |
| median_value         | Median value of the period           | double    | \*             |
| max_value            | Max value of the period              | double    | \*             |
| date_max_value       | Date one which max value was reached | double    | \*\*           |
| days_onset_abs       | Number of days between the 1^s^t day | double    | days           |
|                      | and the day of maximal value         |           |                |
| raw_onset_rate_abs   | Slope between 1^s^t day and the day  |           |                |
|                      | of maximal value                     | double    | \* per day     |
| days_offset_abs      | Number of days between the day of max| double    | days           |
|                      | value and the last day               |           |                |
| raw_offset_rate_abs  | Slope between the day of maximal     |           |                |
|                      | value and the last day               | double    | \* per day     |
| daily_rate           | variation between d day and d-1      | double    | \* per day     |
| baseline_qt          | value in baseline_qt argument        | double    | \*             |
| baseline_mean        | value in baseline_mean argument      | double    | \*             |
| anomaly_qt           | value on d day - baseline_qt on d day| double    | \*             |
| anomaly_mean         | value on d day - baseline_mean on d  | double    | \*             |
|                      | day                                  |           |                |
| anomaly_unit         | baseline_qt on d day - baseline_mean | double    | \*             |
|                      | on d day                             |           |                |
| daily_category       | period category according Hodbay *et*| factor (5)| no,I,II,III,IV |
|                      | *al.*, 2018                          |           |                |
| mean_anomaly_qt      | average of anomaly_qt in the period  | double    | \*             |
| sd_anomaly_qt        | standard deviation of anomaly_qt in  | double    | \*             |
|                      | the period                           |           |                |
| max_anomaly_qt       | maximum value of anomaly_qt in the   | double    | \*             |
|                      | period                               |           |                |
| mean_anomaly_mean    | average of anomaly_mean in the period| double    | \*             |
| sd_anomaly_mean      | standard deviation of anomaly_mean in| double    | \*             |
|                      | the period                           |           |                |
| max_anomaly_mean     | maximum value of anomaly_mean in the | double    | \*             |
| mean_anomaly_fixed   |average of anomaly_fixed in the period| double    | \*             |
| sd_anomaly_fixed     | standard deviation of anomaly_fixed  | double    | \*             |
|                      | in the period                        |           |                |
| max_anomaly_fixed    | maximum value of anomaly_fixed in    | double    | \*             |
|                      | the period                           |           |                |
|most_frequent_category| most frequent category in the period | character |                |
| evolution_rate_lag_x | rate of 'value' computed over x days | double    | \* per day     |
|                      | available when time_lapse_vector=x   |           |                |
| variance_value_lag_x | variance of 'value' computed over x  | double    | \* squared     |
|                      |days, available if time_lapse_vector=x|           |                |  

\* Same unit as the spatraster provided (here it is degree Celsius)
\*\* Same format as the in the spatraster provided (here yyyy-mm-dd)  

Note: columns depending on baseline_qt, baseline_mean and baseline_fixed will only be calculated if the associated argument have been provided.  


# Examples
## Load data:
```{r setup}
library(BioExtremeEvent)
file_name_1 <- system.file(file.path("extdata", 
                                      "copernicus_data_celsius.tiff"),
                                      package = "BioExtremeEvent")
 copernicus_data_celsius <- terra::rast(file_name_1)
 file_name_2 <- system.file(file.path("extdata", 
                                      "binarized_corrected_df.rds"),
                                      package = "BioExtremeEvent")
 binarized_corrected_df <- readRDS(file_name_2)
 file_name_3 <- system.file(file.path("extdata", 
                                      "baseline_qt90_smth_15.tiff"),
                                      package = "BioExtremeEvent")
 baseline_qt90_smth_15 <- terra::rast(file_name_3)
 file_name_4 <- system.file(file.path("extdata", 
                                      "baseline_mean_smth_15.tiff"),
                                      package = "BioExtremeEvent")
 baseline_mean_smth_15 <- terra::rast(file_name_4)
 gps <- data.frame(x = c(3.6883,4.4268), # Sète, Saintes-Marie-de-la-Mer
                   y = c(43.3786,43.4279))
```

## Get:

```{r}

```

## Get:
```{r}

```

## Get:

```{r}

```

## Compute metrics for every pixels in yourspatraster:

Apply function:
```{r}

```

## How to save the ouputs:
Outputs are :
```{r, eval = FALSE}

```

# Plots

## Plot:
```{r, fig.width = 20, fig.height =5, warning = FALSE}

```

## Plot:
```{r}


```